---
title: "Make my plot prettier"
author: "Antonio Miguel de Jesus Domingues"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width=7,
  fig.height=7,
  comment = "#>"
)
```

This vignette will introduce you to the package `makeitprettier`, which is fairly mundane affair, but will also be used to collect the tips and tricks that I frequently use when building plots/figures for presentation/publication and that could not be wrapped up in a function. 


## Installation

The easiest way is using `devtools` to install straight from github:

```r
# install.packages("devtools")
devtools::install_github("adomingues/makeitprettier")
```


## Usage

The package is organized in functions that can be added straighforwardly to any plot. Those functions are define as: 

- theme_* to change the appearance of the plot
- scale_color_* and scale_fill_ to change the colors used in the plot.

I kept the theme simple and clean so that we can focus on the data, but did not get rid of all gridlines because I find those helpful to read values in a plot, ins particular in a poster when people may actually want to look at the detail.

So how to actually use it? I have included a data set that mimics a typical IP experiment, with 3 replicates and 3 different antibodies to show what we can do with it. First let's plot the number of reads for each length :


```{r test-data}
library("makeitprettier")
head(readcounts)

p1 <- ggplot(readcounts, aes(x = readlength, y = N, color = Mapping)) +
  geom_line() +
  facet_grid(IP ~ Replicate)
p1
```

To make it prettier we just add our little functions:

```{r pretty}
p1 + 
  theme_poster() +
  scale_color_prettier()
```

We did away with most of the visual clutter to make a much cleaner plot. I am planning on `publication` function that does away with even more visual elements:

- Remove the gray boxes for each panel
- Remove the gridlines


## (brief) ggplot explanation 

The plot above is a fairly complicated visualization in which we want to compare multiple things:

- Sense vs antisense reads, which we can compare via the line color
- Are the replicates similar? Easy to compare horizontally
- Do the three IPs yield the same results?

In `ggplot` we can map a certain visual attribute to a condition/factor. For instance, line color to the "Mapping". We can also use the factors to split the data in rows/columns as we did with `facet_grid(IP ~ Replicate)` where we are simply telling `ggplot` that we can split those variables in a grid with the IPs horizontally and the replicates vertically. 


## Further improvements

You would never put this in a final manuscript! The axis titles are either not readable, not informative or ugly. The values in a the y-axis are also pretty hard to read in a quick glance. Let's fix it.

```{r fix-it}
p2 <- p1 + 
  theme_poster() +
  scale_color_prettier()+
  scale_y_continuous(labels = scales::comma) +
  labs(
    x = "Read length (nt)",
    y = "Number of reads") 

p2
```

There you have it. It could almost go like this into print.


## Save plots

As you probably noticed, by default `gpplot` saves plots which requite resizing before inclusion in a poster. I have included a little function that saves the plots:

- in 15 x 15 cm (the default is inches - yuk)
- as _pdf_ and _png_, for flexibility


```{r save-plot-1}
save_plot("~/test_save")
```

The function knows which plot was generated last and saves that one. Regardless, it can also be told which plot to save:

```{r save-plot-2}
save_plot("~/test_save", p2)
```

I found that 15 x 15 cm was about the right size for figure to slapped straight into a poster. The function takes all the usual `ggsave` arguments so the plot dimensions can be changed at will.

It might (and it will be) useful to save the data in the plot to re-do it at a later stage without doing the analysis all over again. For that there is a little function:


```{r save-data-1}
save_data("~/test_save.data.txt")
save_data("~/test_save.data.txt", p2)
```

The above will do the same, and will this:

```{r save-all}
save_plot("~/test_save", save_data = TRUE)
```